---
title: "Bayesian Data Analysis with BRMS"
author: |
  | Mitzi Morris
  | Stan Development Team
  | Columbia University, New York NY
date: Feb 28, 2023
output:
  html_document:
    toc: true
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=FALSE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='90%')
```

## Getting Started

Install, load packages and data used in this notebook

```{r}
packages <- c("readr", "ggplot2", "brms", "bayesplot", "loo", "projpred", "cmdstanr")
require(packages)
suppressMessages(  lapply(packages, library, character.only = TRUE)  )
options(brms.backend = "cmdstanr")
```

The data and models are taken from chapter 12 of the book [Data Analysis
Using Regression and Multilevel/Hierarchical
Models](http://www.stat.columbia.edu/~gelman/arm/) by Andrew Gelman and
Jennifer Hill, Cambridge Press, 2006.

## Example: EPA Measurements of Radon Levels in the Home

This example is taken from chapter 12 of the book [Data Analysis Using
Regression and Multilevel/Hierarchical
Models](http://www.stat.columbia.edu/~gelman/arm/) by Andrew Gelman and
Jennifer Hill, Cambridge Press, 2006.

In the early 1990s, the US Environmental Protection Agency (EPA) did a
national survey of home radon levels. Gelman and Hill analyze data for
the state of Minnesota.

-   Survey data
    -   919 measurements of radon gas levels in residential houses
    -   85 counties in Minnesota have survey data, 2 counties have no
        data.
    -   measurements were taken in the basement, if any, else on the
        ground floor.
    -   another EPA dataset provides the soil uranium level for each
        county
-   Regression model
    -   outcome: home radon level (on the log scale)
    -   predictors: county, floor, uranium (on the log scale)

The data file is `data/mn_radon.csv` and the data is loaded into
dataframe `mn_radon`.

```{r}
mn_radon <- read_csv("data/mn_radon.csv")
mn_radon$county_id <- as.factor(mn_radon$county_id)
print(mn_radon)
```

## Modeling Choices

Structure in the data: houses can be grouped by county.

-   Modeling choices
    -   Complete pooling: all counties are alike. Run a single
        regression to estimate average home radon level in MN.
    -   No pooling: all 85 counties are different. Run 85 separate
        regressions to estimate home radon levels by county.
    -   Partial pooling: counties are similar. Build a multi-level
        regression to share information across counties.

## Complete Pooling Model

The simplest model is the complete pooling model which estimates the
home log radon level based on the floor on which the measurement was
taken.

The complete pooling formula looks just like a simple linear regression
formula

    log_radon ~ floor

We plug this into the `brm` function and summarize the fit.

```{r}
fit_complete_pool = brm(log_radon ~ floor, data=mn_radon)
summary(fit_complete_pool)
```

The function `conditional_effects` plots the outcome on the y axis and
the predictor on the x axis.

```{r}
conditional_effects(fit_complete_pool)
```

## No Pooling Model

The no pooling formula estimates individual per-county intercepts, each
intercept is estimated separately.

    log_radon ~ 0 + floor + county_id

```{r}
fit_no_pool = brm(log_radon ~ 0 + floor + county_id, data=mn_radon)
summary(fit_no_pool)
conditional_effects(fit_no_pool, effects = c("floor:county_id"))
```

## Partial Pooling Model: varying slope

The partial pooling formula estimates per-county intercepts which drawn
from the same distribution which is estimated jointly with the rest of
the model parameters. The `1` is the intercept co-efficient. The
estimates across counties will all have the same slope.

    log_radon ~ floor + (1 | county_id)

```{r}
fit_part_pool_1 = brm(log_radon ~ floor + (1 | county_id), data=mn_radon)
summary(fit_part_pool_1)
bayesplot::mcmc_intervals(fit_part_pool_1, regex_pars = "r_county") +  bayesplot::yaxis_text(FALSE)
```

## Partial Pooling Model: varying slope, varying intercept

The varying-slope, varying intercept model adds `floor` to the
group-level co-efficients. Now estimates across counties will all have
varying slope.

    log_radon ~ floor + (1 + floor | county_id)

```{r}
fit_part_pool_2 = brm(log_radon ~ floor + (1 + floor | county_id), data=mn_radon)
bayesplot::mcmc_intervals(fit_part_pool_2, regex_pars = "r_county") +  bayesplot::yaxis_text(FALSE)
```

## Partial Pooling Model: varying slope, varying intercept, another predictor

```{r}
fit_part_pool_3 = brm(log_radon ~ floor + log_uranium + (1 + floor | county_id), data=mn_radon)
bayesplot::mcmc_intervals(fit_part_pool_3, regex_pars = "r_county") +  bayesplot::yaxis_text(FALSE)
```
